---
title: "Introduction"
author: "Nate Lant"
date: "3/31/2020"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(nhts2017)
library(scales)

```

## The National Household Travel Survey
The National Household Travel Survey (NHTS) is conducted every 8-10 years or so,
and samples nearly 300,000 americans to understand travel behavior. Each person
answers questions about themselves and their household (i.e. employment status, 
income, age, gender
etc.) and records information on their travel behavior each day of the study.

Each respondant is asked if they have any kind of disability that limits their
transportation. If they respond yes, then they are asked which kinds of medical 
devices they use (i.e. wheelchair, cane, seeing eye dog, etc.). From these 
responses, we can label each person as wheelchair dependent or not.

```{r data_cleaning, eval=F}
mydata <- nhts_persons %>%
  select(houseid, personid, r_age, educ, r_sex, r_race, wtperfin, medcond,
         w_chair, w_scootr, w_mtrchr, worker, wrk_home, wkftpt) %>%
  left_join(
    nhts_households %>% select(houseid, msasize, hhfaminc),
    by = "houseid"
  ) %>%
  filter(r_age > 24 & r_age < 65) %>%
  mutate(
    worker = ifelse(worker == "01", 1, 0),
    wheelchair = ifelse(
      w_chair == "07" | w_mtrchr == "08" | w_scootr == "06", 1, 0
    ),
    disability = ifelse(medcond == "01", 1, 0)
  ) %>%
  select(-medcond, -w_chair, -w_scootr, -w_mtrchr) %>%
  filter(!educ %in% c("-7", "-8", "-1"),
         !r_sex %in% c("-7", "-8"),
         !hhfaminc %in% c("-7", "-8", "-9"),
         !r_race %in% c("-7", "-8")) %>%
  mutate(
    inc_cont = case_when(
      hhfaminc == "01" ~ 5000,
      hhfaminc == "02" ~ 12500,
      hhfaminc == "03" ~ 20000,
      hhfaminc == "04" ~ 30000,
      hhfaminc == "05" ~ 42500,
      hhfaminc == "06" ~ 62500,
      hhfaminc == "07" ~ 87500,
      hhfaminc == "08" ~ 112500,
      hhfaminc == "09" ~ 137500,
      hhfaminc == "10" ~ 175000,
      hhfaminc == "11" ~ 200000,
    ),
    male = ifelse(r_sex == "01", 1, 0),
    college_plus = ifelse(educ %in% c("03", "04", "05"), 1, 0),
    pop_million = ifelse(msasize %in% c("04", "05"), 1, 0),
    wrk_home = ifelse(wrk_home == "01", 1, 0)
  )

mydata %>% write_csv("data/updated.csv")
mydata %>% write_rds("data/mydata.rds")
```

## Wheelchair Status and Labor Force Participation
This study will evaluate the effect of being limited to a wheelchair on participation in
the labor force. As shown in the figure below, there is a major imbalance in the number of
individuals limited to a wheelchair who are employed versus unemployed, 
but could that be because of their age or because of their educational 
attainment level, instead of their dependence to a wheelchair?

Or we could ask, does being limited to a wheelchair affect employment more than simply
being disabled?

```{r, message=F, warning=F}
# show plot of distribution (worker and wheelchair)
mydata <- read_rds("data/mydata.rds")

mydata %>%
  group_by(worker, wheelchair) %>%
  summarise(population = sum(wtperfin)) %>%
  group_by(wheelchair) %>%
  mutate(distribution = population/sum(population)) %>%
  ggplot(aes(x = as.character(wheelchair), y = distribution)) +
  geom_col(aes(fill = as.character(worker)), colour = "Black", position = "dodge") +
    ggtitle("Distribution of Worker Status and Wheelchair Usage") +
    labs(x = "Wheelchair Usage (1 = yes)", y = "Distribution") +
    scale_fill_brewer(palette = "PuBuGn", direction = 2) + 
    theme(axis.text.x = 
            element_text(size  = 10, angle = 45, hjust = 1, vjust = 1))
```

It is obvious that by having a wheelchair, liklihood of unemployment is much greater. However, 
other variables may be the cause of unemployment. Below is a summary table of the other variables 
we will consider in our model.

```{r summary}
mydata %>%
  group_by(worker, wheelchair) %>%
  summarise(respondents = n(),
            count = sum(wtperfin),
            av_age = mean(r_age),
            median_inc = median(inc_cont),
            male = mean(male),
            wrk_home = mean(wrk_home),
            college_plus = mean(college_plus),
            pop_million = mean(pop_million)) %>%
  ungroup() %>%
  mutate(dist = percent(count/sum(count), accuracy = 0.1)) %>%
  select(-count)
```



```{r}
mydata %>%
  group_by(wheelchair, worker)%>%
  tally() %>%
  spread(wheelchair, n)

mydata %>%
  filter(worker == 1) %>%
  ggplot(aes(r_age, fill = as_factor(wheelchair))) + geom_density(alpha = 0.2)
```



