---
title: "Effect of Wheelchair Use"
author: "Nate Lant and Austin Hopkins"
date: "4/14/2020"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

mydata1 <- read_rds("data/mydata.rds") %>%
  mutate(age_bar = ifelse(r_age > 55, 1, 0))
```

This section compares different models in predicting the likelihood of employment for users
with wheelchairs. The response variable is the `worker` variable which is employment status, 0 
for unemployed and 1 for employed.



At the most basic level, the binomial regression is the same as the odds ratio. The logit is $z$ in 
$$  z = \beta_0 + \beta_1*x_1 $$

The odds ratio can be calculated using $e^\beta$. This is not to be confused with the odds,
defined as $e^z$. With odds, we can calculate the probability of getting employed as

$$ P = \frac{e^z}{1+e^z}$$

## Reduced Model

The first model will regress worker status on wheelchair usage. This model will calculate the 
odds of employment if you have a wheelchair compared to not having a wheelchair (odds ratio).

```{r reduced}
glm(worker ~ wheelchair, data = mydata1, family = binomial(link='logit')) %>%
  summary()
```

In this case, $e^(-2.976)$ is 0.051, or 1:20. This states that if you have a wheelchair, you 
are 20 times more likely to be unemployed. Lets see how this changes with interactions.


## Full Model

```{r full}
glm(worker ~ educ*wheelchair + r_sex*wheelchair + log(inc_cont)*wheelchair + age_bar*wheelchair, 
    data = mydata1, family = "binomial") %>%
  summary()
```

This shows for example the odds ratio of a wheelchair user who attained education level 5
(graduate or professional degree) is $e^(-3.03+1.03+1.175)$ = `r exp(-3.03+1.03+1.175)%>%round(3)`.

This shows that education (with the exception of graduate school) does not affect employment for
wheelchair users any differently than for non wheelchair users.






